\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
Chapter 1 (add link) talks about ...

In Chapter 2 we discuss ... This work resulted in the publication of the following article:
\begin{quote}
William McCorkindale, Ivan Ahel, Haim Barr, Galen J. Correy, James S. Fraser, Nir London, Marion Schuller, Khriesto Shurrush, Alpha A. Lee. Fragment-Based Hit Discovery via Unsupervised Learning of Fragment-Protein Complexes.
\end{quote}
Specify who did what

Chapter 3:
\begin{quote}
Aaron Morris, William McCorkindale, The COVID Moonshot Consortium, Nir Drayman, John D. Chodera, Savaş Tay, Nir London and Alpha A. Lee. Discovery of SARS-CoV-2 main protease inhibitors using a synthesis-directed de novo design model, \textit{Chem. Commun.}, 2021,57, 5909-5912 
\end{quote}

Chapter 4
\begin{quote}
Dávid Péter Kovács, William McCorkindale and Alpha A. Lee. Quantitative interpretation explains machine learning models for chemical reaction prediction and uncovers bias. \textit{Nature Communications} volume 12, Article number: 1695 (2021)
\end{quote}

D.P.K. and W.M. implemented the algorithms, D.P.K. trained the models, designed the experiments and analysed the model attributions for the various reaction classes. W.M. implemented Tanimoto splitting and applied reaction templates for counting statistics and artificial dataset generation. A.A.L. supervised and directed the project. All authors discussed the results and approved the manuscript. D.P.K. and W.M. contributed equally to this study.

Chapter 5:

The final chapter ...

